---
title: "6-2Way ANOVA"
author: "C.M. Gienger"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(ggfortify)
library(emmeans)
library(multcomp)
library(multcompView)
```

```{r include=FALSE, echo=FALSE}
#change depending on whether you want just questions or questions AND answers
knitr::opts_chunk$set(fig.path='Figs/', echo=FALSE, warning=FALSE, message=FALSE)
```
```{r}
#http://r4all.org/posts/importing-data-update/
#can fix by using string as factors option (button)
```


```{r plant growth}
growth.moo <- read.csv("growth.csv")
```

```{r}
glimpse(growth.moo)
#see character data as factors
```
```{r}
growth.moo$diet <-as.factor(growth.moo$diet)
levels(growth.moo$diet)
```
```{r}
growth.moo$supplement <-as.factor(growth.moo$supplement)
levels(growth.moo$supplement)
```

```{r}
# relevel the supplement column
growth.moo <- mutate(growth.moo,
supplement = relevel(supplement, ref="control"))
```

```{r}
# check it worked
levels(growth.moo$supplement)
```

```{r}
# calculate mean and sd of gain for all 12 combinations
sumMoo <- growth.moo %>%
group_by(diet, supplement) %>%
summarise(meanGrow = mean(gain))

# make sure it worked
sumMoo
```
```{r}
#as always, make a plot first
ggplot(sumMoo, aes(x = supplement, y = meanGrow)) +
  geom_point() +
  theme_bw()
```
```{r}
ggplot(sumMoo, aes(x = supplement, y = meanGrow,
colour = diet, group = diet)) +
geom_point() +
geom_line() +
theme_bw()
```
```{r message=FALSE, warning=FALSE}
model_cow <- lm(gain ~ diet * supplement, data = growth.moo)
#examine model
autoplot(model_cow, smooth.colour = NA)
```
```{r}
anova(model_cow)
```
```{r}
summary(model_cow)
```
```{r message=FALSE, warning=FALSE}
# calculate mean and se of gain for all 12 combinations
sumMoo <- growth.moo %>%
  group_by(diet, supplement) %>%
  summarise(
    meanGrow = mean(gain),
    seGrow = sd(gain)/sqrt(n())
)
sumMoo
```
```{r}
ggplot(sumMoo, aes(x = supplement, y = meanGrow,
                   colour = diet, group = diet)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = meanGrow - seGrow,
                    ymax = meanGrow + seGrow), width = 0.1) +
theme_bw()
```
**CO2 Example**
The CO2 uptake of plants from Quebec and Mississippi was measured at several levels of ambient CO2 concentration. Half the plants of each type were chilled overnight before the experiment was conducted.

Source

Potvin, C., Lechowicz, M. J. and Tardif, S. (1990) “The statistical analysis of ecophysiological response curves obtained from experiments involving repeated measures”, Ecology, 71, 1389–1400.

Pinheiro, J. C. and Bates, D. M. (2000) Mixed-effects Models in S and S-PLUS, Springer.
```{r}
data_CO2<-CO2
```

```{r}
summary_CO2<-data_CO2 %>%
  group_by(Type, Treatment) %>%
  summarise(
    mean_uptake=mean(uptake),
    n=n(),
    se_uptake=sd(uptake)/sqrt(n)
  )
summary_CO2
```

```{r}
plot_CO2 <- ggplot(summary_CO2, aes(x=Type, y=mean_uptake, fill=Treatment)) +
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar(aes(ymin=mean_uptake-se_uptake, ymax=mean_uptake+se_uptake), width=.2, position=position_dodge(.9))+
  ylab("CO2 uptake rates (umol/m^2 sec)")+
  theme_bw()

plot_CO2
```
Making this figure is not as intuitive as you might think because of the position dodge required, 
the stat="identity", and error bar specification.
Really messes up things; not an intuitive intro example.
Note that the reference group in the model summary table is the mean of Quebec non-chilled. Everything else is in reference to that.

```{r}
model1 <- lm(uptake~Type*Treatment, data=data_CO2)
anova(model1)
```

```{r}
summary(model1)
```
```{r}
#Multiple Contrasts using EMmeans package
emm_model1<- emmeans(model1, "Type", "Treatment")
emm_model1
pairs(emm_model1)
```

```{r}
#visually compare marginal means (LS means)
plot(emm_model1, comparisons = TRUE)+
  coord_flip()
```
Because there are two treatments, it breaks them into two panels.
```{r}
#connecting letters differences report
multcomp::cld(emm_model1)
```
Simple Contrasts
```{r}
contrast(emm_model1, "consec", simple = "each", combine = TRUE, adjust = "mvt")
```
Pairwise P-value plots; pwpp plots
```{r}
#a nuanced approach to multiple comparisons
model.cells <- emmeans(model1, ~ Type * Treatment)
pwpp(model.cells, type = "response")
```
```{r}
#add CLD notation from multcomp::cld
#non-nuanced approach
plot_CO2 +
  annotate("text", x = .75, y = 40, label = "2")+
  annotate("text", x = 1.25, y = 35, label = "2")+
  annotate("text", x = 1.75, y = 30, label = "1")+
  annotate("text", x = 2.25, y = 20, label = "1")+
  annotate("text", x = 2, y = 37.5, label = "Shared letters indicate no pairwise \n differences  (P > 0.05) between groups ")
```





